Iterative Monomorphisation

### 2024-02-20 iterative_monomorphisation.tex
continued working on high level description, particularly the example

\begin{definition}
Two substitutions \(\{\alpha_1 \mapsto \tau_1, \dots, \alpha_n\mapsto\tau_n\}\) and \(\{\beta_1 \mapsto \upsilon_1, \dots, \beta_m\mapsto\upsilon_m\}\) are \emph{compatible} if for all \(i, j\): \(\alpha_i = \beta_j\) implies \(\tau_i = \upsilon_j\). 
\end{definition}

@@

\subsection{Collection}
The initial phase of each iteration consists in collecting the elements of the maps \(M\) and \(P\), defined as follows:
\begin{itemize}
    \item Given a type \(n\)-ary symbol \(f\) in \(\Phi\). \(M(f)\) is the set of all \(n\)-tuples of monomorphic type arguments to which \(f\) is applied in \(\Phi\).

    For example, if \(f\langle \tau_1, \tau_2, \tau_3\rangle\) (with \(\tau_1, \tau_2\) and \( \tau_3\) monomorphic) occurs in \(\Phi\) then, \((\tau_1, \tau_2, \tau_3) \in M(f) \).

 \item Given a formula \(\varphi \in \Phi\) and a type \(n\)-ary symbol \(f\) in \(\varphi\). \(P(\varphi, f)\) is the set of all \(n\)-tuples of type arguments to which \(f\) is applied in \(\varphi\). At least one of these type arguments must be polymorphic.

    For example, if \(f\langle \tau_1, \tau_2, \tau_3\rangle\) (with at least one of \(\tau_1, \tau_2\) or \( \tau_3\) polymorphic) occurs in \(\varphi\) then, \((\tau_1, \tau_2, \tau_3) \in P(\varphi, f) \).

@@

\subsection{Iteration}
A single iteration will consist in the following:
\begin{itemize}

@@

      An empty set of substitutions \(\Sigma_{\varphi, f}\) is created, then, for each type argument tuple \((\tau_1, \dots, \tau_n) \in  P(\varphi, f)\) and each type argument tuple \((\upsilon_1, \dots, \upsilon_n) \in M(f)\):

@@

After a set number of iterations, all monomorphic formulae are collected from \(\Phi\) and output.

@@

Consider a problem made up of the following formulae:
\begin{enumerate}
   \item \(\forall a: \alpha, b:\ty{list}(\alpha), f\langle\alpha,\ty{list}(\alpha)\rangle(a, b)\)
   \item \(\forall i: \ty{int}, j: \ty{list}(\ty{int}), f\langle \ty{int}, \ty{list}(\ty{int})\rangle(i, j)\)
\end{enumerate}

>>>>

\begin{definition}
Two substitutions \(\{\alpha_1 \mapsto \tau_1, \dots, \alpha_n\mapsto\tau_n\}\) and \(\{\beta_1 \mapsto \upsilon_1, \dots, \beta_m\mapsto\upsilon_m\}\) are \emph{compatible} if for all \(i, j\), \(\alpha_i = \beta_j\) implies \(\tau_i = \upsilon_j\). 
\end{definition}

@@

\subsection{Collection}
The initial phase of each iteration consists in collecting the elements of the maps \(M\) and \(P\).

Given a type \(n\)-ary symbol \(f\) in \(\Phi\). \(M(f)\) is the set of all \(n\)-tuples of monomorphic type arguments to which \(f\) is applied in \(\Phi\). For example, if \(f\langle \tau_1, \tau_2, \tau_3\rangle\) (with \(\tau_1, \tau_2\) and \( \tau_3\) monomorphic) occurs in \(\Phi\) then, \((\tau_1, \tau_2, \tau_3) \in M(f) \).

Given a formula \(\varphi \in \Phi\) and a type \(n\)-ary symbol \(f\) in \(\varphi\). \(P(\varphi, f)\) is the set of all remaining \(n\)-tuples of type arguments to which \(f\) is applied in \(\varphi\) which are not in \(M(f)\). Meaning that at least one the type arguments is polymorphic. For example, if \(f\langle \tau_1, \tau_2, \tau_3\rangle\) (with at least one of \(\tau_1, \tau_2\) or \( \tau_3\) polymorphic) occurs in \(\varphi\) then, \((\tau_1, \tau_2, \tau_3) \in P(\varphi, f) \).

@@

\subsection{Iteration}
Each iteration of the algorithm follows these steps:
\begin{itemize}

@@

      An empty set of substitutions \(\Sigma_{\varphi, f}\) is created, then, for every type argument tuple \((\tau_1, \dots, \tau_n) \in  P(\varphi, f)\) and every type argument tuple \((\upsilon_1, \dots, \upsilon_n) \in M(f)\):

@@

After a set number of iterations, all monomorphic formulae in \(\Phi\) are output.

@@

Consider a problem made up of the following formulae:
\begin{enumerate}
   \item \(\forall a: \alpha, b:\ty{list}(\alpha), f\langle\alpha\rangle(a) \land f\langle \ty{list}(\alpha)\rangle(b)\)
   \item \(\forall i: \ty{int}, f\langle \ty{int}\rangle(i)\)
\end{enumerate}

\begin{table}[ht]
\caption{State of the algorithm}
\centering\begin{tabular}{@{}lccccccc@{}}
   \toprule
               & Initialisation & Iteration 1 & Iteration 2 & Iteration 3\\
   \midrule
   \(M\)       & \((\ty{int})\)                     & -                     & \((\ty{list}(\ty{int}))\) \\
   \(P(1, f)\) & \((\alpha)\; (\ty{list}(\alpha))\) & -                     & - \\
   \(P(2, f)\) & \(\emptyset\)                      & -                     & - & \(\emptyset\) \\
   \(\Phi\)    & [1] \; [2]                         & \(\sigma(\text{[1]})\)& \(\sigma'(\text{[1]})\)& \\
   \bottomrule
\end{tabular}
\end{table}

On the first iteration, two matchings are attempted, that of \(\alpha\) against \(\ty{int}\) which yields \(\sigma = \{\alpha \mapsto \ty{int}\}\) and that of \(\ty{list}(\alpha)\) against \(\ty{int}\) which fails. Then, \(\sigma\) is applied to [1], resulting in a new formula: \(\forall a: \ty{int}, b:\ty{list}(\ty{int}), f\langle\ty{int}\rangle(a) \land f\langle \ty{list}(\ty{int})\rangle(b)\)

At the beginning of the second iteration, the clause generated in the previous iteration is used to update \(M\).
